Cage effect

### 2018-11-19 cage_effect.Rmd
T7 no 24/7

---
title: "GG_topi2"
author: "Francesco Beghini"
date: "26/09/2018"
output: 
  html_document: 
    theme: cerulean
editor_options: 
  chunk_output_type: console
---

@@

sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])

sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))

@@

rownames(tax_table_filtered) <- rownames(tax_table(gg_topi))
tax_table(gg_topi) <-  as.matrix(tax_table_filtered)

gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))

gg_topi.phylum <- tax_glom(gg_topi,taxrank = "Phylum")
gg_topi.phylum.relab <- transform_sample_counts(gg_topi.phylum, function(x) x/sum(x))

meta_samples <- data.frame(otu_table(gg_topi), check.names = FALSE) %>% 
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  tibble::add_column(., timepoint = stringr::str_sub(.$SampleID, 12, 13)) %>%
  dplyr::group_by(timepoint, OTU) %>%
  dplyr::select(-SampleID)%>%
  dplyr::summarise(Abundance=mean(Abundance)) %>%
  tidyr::spread(timepoint, Abundance) %>%
  tibble::column_to_rownames(var = 'OTU') %>%
  data.frame()

gg_topi_meta_samples <- phyloseq(otu_table(meta_samples, taxa_are_rows = TRUE), tax_table(gg_topi), phy_tree(gg_topi))
```
<!-- # Taxonomy profile at genus level -->
<!-- ```{r taxonomy, cache=FALSE} -->
<!-- # nychanesmicrobiome::plot_abundance(gg_topi,taxrank = 'Genus') -->
<!-- ``` -->

VVP: Vuote + PBS

VPP: Pentatoma + PBS

VVB: Vuote + *Bifidobacterium*

VPB: Pentatoma + *Bifidobacterium*

# Alpha diversity
```{r Alpha diversity}
# otu_table(gg_topi) %>% colSums() %>% sort %>% barplot(xlab = c())
# otu_table(gg_topi) %>% colSums() %>% summary()

alpha_div <- estimate_richness(gg_topi)
alpha_div %<>% bind_cols(data.frame(sample_data(gg_topi)))

x <- melt(alpha_div, id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon'))
x$value <- as.numeric(x$value)

ggplot(x) +
  geom_boxplot(aes(timepoint, value)) +
  facet_grid(variable~condizione,scales = 'free_y')
```
<!-- ## SFB abundance -->
<!-- ```{r} -->
<!-- knitr::kable(otu_table(gg_topi)[c(879,880),]) -->
<!-- ``` -->

# Beta diversity
```{r Beta diversity, cache=TRUE}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
distbc <- phyloseq::distance(gg_topi, "bray")
distjac <- phyloseq::distance(gg_topi, "jaccard")
distwu_meta <- phyloseq::distance(gg_topi_meta_samples, "jaccard")

ordwu <-  ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
ordwu_meta <-  ordinate(gg_topi_meta_samples, method = "MDS", distance = distwu_meta, weighted = TRUE)
ordbc <-  ordinate(gg_topi, method = "MDS", distance = distbc, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
ordjac <-  ordinate(gg_topi, method = "MDS", distance = distjac, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)

# plot_ordination(gg_topi, ordwu, color = 'condizione',shape = 'timepoint', title = 'Weighted UniFrac')
# plot_ordination(gg_topi, ordwu, color = 'merge_timepoint', title = 'Weighted UniFrac')

distances <- reshape2::melt(as.matrix(distwu), varnames = c('rows','cols'))
distances$condizione1 <- sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][2])
distances$pentatome1 <- sapply(as.character(distances$rows), function(x) str_sub(x, 5,5))
distances$condizione2 <- sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][2])
distances$topo1 <- factor(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][3]))
distances$topo2 <- factor(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][3]))
distances$timepoint1 <- stringr::str_replace(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distances$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric

distancesbc <- reshape2::melt(as.matrix(distbc), varnames = c('rows','cols'))
distancesbc$condizione1 <- sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][2])
distancesbc$condizione2 <- sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][2])
distancesbc$topo1 <- factor(sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][3]))
distancesbc$topo2 <- factor(sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][3]))
distancesbc$timepoint1 <- stringr::str_replace(sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distancesbc$timepoint2 <- stringr::str_replace(sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric

distancesjac <- reshape2::melt(as.matrix(distjac), varnames = c('rows','cols'))
distancesjac$condizione1 <- sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][2])
distancesjac$condizione2 <- sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][2])
distancesjac$topo1 <- factor(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][3]))
distancesjac$topo2 <- factor(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][3]))
distancesjac$timepoint1 <- stringr::str_replace(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distancesjac$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
```

## Ordination: Mean all samples per timepoint
```{r}
x <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)

centroids <- aggregate(cbind(Axis.1, Axis.2)~timepoint, x, mean)

ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
  geom_point(size = 0.8) +
  geom_point(data=centroids, size = 4)
```

## Ordination: T2 vs T3, tumor vs no tumor
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T2','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T2','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T2','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=tumor_injected), size=3, alpha=.8) +
  scale_color_brewer(name = "Tumor injected", palette="Dark2") + 
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = -0.2, xmax = -.2, ymin = .25, ymax = .25)
```

## Alpha diversity: tumor vs no tumor
```{r}
filt_alpha <- alpha_div %>% 
  filter(timepoint %in% c('T2','T3')) %>%
  select('Description','condizione','timepoint','Observed', 'se.chao1', 'Simpson', 'Shannon') %>%
  mutate(t=as.numeric(stringr::str_sub(timepoint,2,2)))

ttest_res <- data.frame()
for(measure in c('Observed', 'se.chao1', 'Simpson', 'Shannon')){
  ttest_res %<>% bind_rows(data.frame(Measure=measure, pvalue=with(filt_alpha, t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
}

filt_alpha %>%
  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon')) %>%
  ggplot() +
  geom_boxplot(aes(timepoint, value)) +
  facet_grid(variable~.,scales = 'free_y')

knitr::kable(ttest_res)
```

### Merge T1 and T2 and use them as baseline
## Ordination: baseline vs T3
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T3'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T4
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T4'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T5
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T5'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T6
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T6'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T7
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T7'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: T3 vs T7: Treated with Bifidum?
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('T3','T7'))) +
  geom_point(aes(Axis.1, Axis.2, color=cond_all), size=3) +
  scale_colour_discrete(name="Timepoint_Vaccine_Bifido") +
  theme_bw() 
```

```{r}
ggplot(subset(ordwu, timepoint %in% c('T3','T7'))) +
  geom_point(aes(Axis.1, Axis.2, color=timepoint, shape=bifido), size=3) +
  facet_wrap(pentatome~bifido)+
  theme_bw() 
```

## Ordination: T3 vs T4
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T4','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T4','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ timepoint + real_vaccine + timepoint:real_vaccine, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T4','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint, shape=real_vaccine), size=3, alpha=.7) +
  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
  scale_shape_manual(name = 'Pentatome?', values=c(15,19))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('Timepoint PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = .2, xmax = .2, ymin = .1, ymax = .1) +
  annotation_custom(grob = grid::textGrob(paste('Pentatome PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[2], '\nR2=', signif(permanova$aov.tab$R2[2], 2)), 
gp=grid::gpar(col="black", fontsize=10)),  xmin = -.1, xmax = -.1, ymin = -.08, ymax = -.08)
```

## Ordination: T3 vs T5
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T5','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T5','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T5','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .12, ymax = .12)
```

## Ordination: T3 vs T6
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T6','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T6','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T6','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
```

## Ordination: T3 vs T7
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T7','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
```

## Beta diversity between timepoints, T1 is the baseline
```{r Beta diversity between timepoints t1}
distances[distances$timepoint1 == 1 | distances$timepoint1 == 2, 'timepoint1'] <- 2
distances[distances$timepoint2 == 1 | distances$timepoint2 == 2, 'timepoint2'] <- 2

t1 <- distances %>% 
  filter(topo1==topo2) %>%
  filter(timepoint1==2) %>%
  filter(timepoint1!=timepoint2) %>%
  arrange(topo1)

t.res <- data.frame()
for(cond in unique(t1$condizione1)){
  for(x in combn(c(3:7), 2, simplify = FALSE)){
    pval.t = t.test(t1[t1$timepoint2==x[1] & t1$condizione1==cond,"value"], t1[t1$timepoint2==x[2] & t1$condizione1==cond ,"value"])$p.value
    pval.w = wilcox.test(t1[t1$timepoint2==x[1] & t1$condizione1==cond,"value"], t1[t1$timepoint2==x[2] & t1$condizione1==cond ,"value"], paired = F)$p.value
    if(pval.t < 0.05 | pval.w < 0.05)
      t.res %<>% bind_rows(c(condizione1=cond, timepoint=as.numeric(x), p.value = str_c('t-test:',signif(pval.t, digits = 2),'U-test:', signif(pval.w, digits = 2), sep = ' ')))
    }
  }

```{r}
knitr::kable(t.res)
```

```{r}

t1 %>%
  ggplot(aes(timepoint2, value, group = timepoint2)) +
  stat_boxplot() +
  scale_x_continuous(breaks = c(2:7)) +
  # geom_signif(data=t.res,
              # aes(xmin= as.numeric(timepoint1), xmax = as.numeric(timepoint2), annotations = p.value, y_position = 0.33),
              # manual = TRUE) +
  facet_grid(condizione1 ~ .)

## Beta diversity between timepoints, T(-1) is the baseline
```{r Beta diversity between timepoints tn-1}
tn1 <- distances %>% 
  filter(topo1==topo2) %>%
  filter(timepoint1==timepoint2-1) %>%
  filter(timepoint1!=timepoint2) %>%
  arrange(topo1)

t.res <- data.frame()
for(cond in unique(tn1$condizione1)){
  for(x in combn(c(3:7), 2, simplify = FALSE)){
      pval.t = t.test(tn1[tn1$timepoint2==x[1] & tn1$condizione1==cond,"value"], tn1[tn1$timepoint2==x[2] & tn1$condizione1==cond ,"value"])$p.value
      pval.w = wilcox.test(tn1[tn1$timepoint2==x[1] & tn1$condizione1==cond,"value"], tn1[tn1$timepoint2==x[2] & tn1$condizione1==cond ,"value"], paired = F)$p.value
      if(pval.t < 0.05 | pval.w < 0.05)
        t.res %<>% bind_rows(c(condizione1=cond, timepoint=as.numeric(x), p.value = str_c('t-test:',signif(pval.t, digits = 2),'U-test:', signif(pval.w, digits = 2), sep = ' ')))
  }
}
```

```{r}
knitr::kable(t.res)
```

```{r}
tn1 %>%
  ggplot(aes(timepoint2, value, group = timepoint2)) +
  geom_boxplot() +
  scale_x_continuous(breaks = c(2:7)) +
  # geom_signif(data=t.res,
              # aes(xmin= as.numeric(timepoint1), xmax = as.numeric(timepoint2), annotations = p.value, y_position = 0.33),
              # manual = TRUE) +
  facet_grid(condizione1 ~ .)
```

## Differential analysis
```{r DA, cache=TRUE}
threshold <- 0.05

ggtopi_filt <- filter_taxa(gg_topi, function(x) mean(x) > 5e-1, TRUE)

dds <- DESeq2::DESeq(phyloseq_to_deseq2(ggtopi_filt, design = ~ cond_vac), parallel = TRUE, quiet = TRUE, fitType = 'parametric')

mmtx <- model.matrix(object = ~timepoint*real_vaccine*real_bifido, data = data.frame(gg_topi@sam_data))
idx <- which(apply(mmtx, 2, function(x) all(x==0)))
mmtx <- mmtx[,-idx]

rmmtx <- model.matrix(object = ~ timepoint, data = data.frame(gg_topi@sam_data))

pds <- phyloseq_to_deseq2(ggtopi_filt, design = ~timepoint*real_vaccine*real_bifido, ignoreRank = TRUE)

dds <- DESeq2::DESeq(pds, full = mmtx, reduced = rmmtx, test = 'LRT')

dds_bif <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi, design = ~ cond_bif), parallel = TRUE, quiet = TRUE)
dds1 <- DESeq2::DESeq(phyloseq_to_deseq2(ggtopi_filt, design = ~ timepoint), parallel = TRUE, quiet = TRUE)

dds_phylum <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi.phylum, design = ~ timepoint), parallel = TRUE, quiet = TRUE)

```
## T2 vs T3
```{r DA T2 vs T3}
res23 <- DESeq2::results(dds1, contrast = c("timepoint","T2","T3"), alpha = 0.05, format = "DataFrame")
res23.filtered <- res23[!is.na(res23$padj),]
res23.filtered <- res23.filtered[res23.filtered$padj < threshold,]
res23.filtered <- res23.filtered[abs(res23.filtered$log2FoldChange) > 2,]

res23.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res23.filtered),])) %>%
  ggplot(aes(log2FoldChange, Genus)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = log(baseMean))) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  ggtitle('T1 vs T2') +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

## T2 vs T3 phylum level
```{r DA T2 vs T3 phylum}
res_phylum <- DESeq2::results(dds_phylum, contrast = c("timepoint","T2","T3"), alpha = 0.05, format = "DataFrame")
res_phylum.filtered <- res_phylum[!is.na(res_phylum$padj),]
res_phylum.filtered <- res_phylum.filtered[res_phylum.filtered$padj < threshold,]
# res_phylum.filtered <- res_phylum.filtered[abs(res_phylum.filtered$log2FoldChange) > 2,]

res_phylum.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res_phylum.filtered),])) %>%
  ggplot(aes(log2FoldChange, Phylum)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

## Check T1 vs T2
```{r DA T1 vs T2}
res12 <- DESeq2::results(dds1, contrast = c("timepoint","T1","T2"))
res12.filtered <- res12[!is.na(res12$padj),]
res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

res12.filtered %>% 
  as.data.frame %>% 
  cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
  ggplot(aes(log2FoldChange, Genus)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  ggtitle('Timepoint 1 vs Timepoint 2') +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

```{r DA T3 vs T4}
res23 <- DESeq2::results(dds1, contrast = c("timepoint","T3","T4"))
res23.filtered <- res23[!is.na(res23$padj),]
res23.filtered <- res23.filtered[res23.filtered$padj < threshold,]

res23.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res23.filtered),])) %>%
  ggplot(aes(log2FoldChange, Genus)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  ggtitle('Timepoint 3 vs Timepoint 4') +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
)
```

## Differential analysis Pentatome T7 vs not vaccinated T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"))
  res12 <- DESeq2::results(dds)
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = log10(baseMean))) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle('Pentatome T3 vs T7') +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

## Differential analysis Pentatome T3 VS T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T3_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = log10(baseMean))) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle('Pentatome T3 vs T7') +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

```{r}
res12.filtered %>% 
  as.data.frame %>% 
  cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
  filter(is.na(Phylum)) %>%  knitr::kable()
```

## Differential analysis Empty vescicles T3 VS T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_NO","T3_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = baseMean)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle('Empty T3vsT7') +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

## Differential analysis Pentatoma vs empty vescicles T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = baseMean)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

## Differential analysis Pentatoma vs empty vescicles T3
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T3_YES","T3_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
  timepoint_number <- stringr::str_sub(timepoint,17,17)

>>>>

---
title: "Cage effect"
author: "Francesco Beghini"
date: "November 8, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])
sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))

@@

rownames(tax_table_filtered) <- rownames(tax_table(gg_topi))
tax_table(gg_topi) <-  as.matrix(tax_table_filtered)
```

# How's the dispostion of cages
```{r Beta diversity, cache=TRUE}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
ordwu <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE)
```

```{r}
plot_ordination(gg_topi, ordwu, color="cage")
```

# B-diversity intra cage
```{r, cache = TRUE}
distances <- reshape2::melt(as.matrix(distwu), varnames = c('rows','cols'))
distances$topo1 <- factor(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][3]))
distances$topo2 <- factor(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][3]))
distances$timepoint1 <- stringr::str_replace(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distances$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric

distances %<>%
  inner_join(unique(data.frame(sample_data(gg_topi))[,c('id.topo','cage')]), by = c('topo1' = 'id.topo')) %>%
  rename('cage'='cage1') %>%
  left_join(unique(data.frame(sample_data(gg_topi))[,c('id.topo','cage')]), by = c('topo2' = 'id.topo')) %>%
  rename('cage'='cage2')
```

# Intra cage diversity through time
```{r}
  distances %>%
  filter(cage1==cage2) %>%
  filter(timepoint1==timepoint2) %>%
  ggplot() +
  geom_boxplot(aes(cage1, value)) + 
  facet_wrap(timepoint1 ~ .)
```

# Inter cage diversity through time
```{r}
  distances %>%
  filter(cage1!=cage2) %>%
  filter(timepoint1==timepoint2) %>%
  mutate(cage = paste(.$cage1, .$cage2, sep =' vs ')) %>%
  ggplot() +
  geom_boxplot(aes(cage, value)) + 
  facet_wrap(timepoint1 ~ ., ncol = 1) +
  theme(axis.text.x = element_text(angle = 90))
```
